---
title: "Geography 176A"
author: "[Liam Finn](https://anthonyfinn.github.io/liam.finn/)"
subtitle: 'Lab 03: COVID-19 Pandemic'
output:
  html_document:
    theme: darkly
---
In this lab, I practiced data wrangling and visualization skills using real-time COVID-19 data maintained by the New York Times.
If you scroll down this page, you will see different tables demonstrating COVID trends across the state of California. 
```{r}
#LOAD DATA
library(tidyverse)
covid_url=("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv")
covid= read_csv(covid_url)  #covid= data from nytimes

```
```{r}
#STEP2: subset Covid Data -> filters data to CA Counties
CA_covid= covid%>%
  filter(state=="California")%>%
  group_by(county) %>%
  mutate(newCases = c(cases[1], diff(cases))) %>%
  ungroup() %>%
  filter(date >=max(date)-13)

#STEP 3
#Most cumulative cases
CA_cumulative= CA_covid %>%
  filter(date==max(date)) %>%
  slice_max(cases, n=5) %>%
  select(county,cases)
#Most new cases
CA_new= CA_covid %>%
  filter(date==max(date)) %>%
  slice_max(newCases, n=5) %>%
  select(county, newCases)

#make table using Knitr::Kable
knitr::kable(CA_cumulative,
             caption = "Counties with Most Cumulative Cases",
             col.names = c("County", "Total Cases"))
knitr::kable(CA_new,
             caption = "Counties with Most New Cases",
             col.names = c("County", "Total New Cases"))

#STEP4/5 : Load population data
```


```{r}
library(readxl)
pop = read_excel('../data/PopulationEstimates.xls', skip = 2) %>%
  select(pop2019 = POP_ESTIMATE_2019, fips = FIPStxt) #change column names

#STEP 6 #examine data
names(pop)
dim(pop)
names(CA_covid)
dim(CA_covid)

#STEP 7
#join COVID & Population data

all_data= left_join(covid, pop, by='fips')


```
These are the 13 counties deemed safe in terms of the number of covid cases among their residents. Over the most recent 14 days, if more than 100 people (per 100,000 residents) are infected with Covid, they are placed on the State Watch List.  
```{r}
last_14_days= all_data %>%
  filter(state==c("California")) %>%
  group_by(county, date) %>%
  summarize(totalcases=sum(cases, na.rm = TRUE),
            pop2019= sum(pop2019, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(date>=max(date)-13)%>%
  group_by(county,pop2019)%>%
  mutate(newCases=totalcases-lag(totalcases))%>%
  summarize(totalcases= sum(newCases,na.rm = TRUE)) %>%
  mutate(per100= totalcases/ (pop2019/100000)) %>%
  filter(per100 <=100)%>%
  pull(county)
last_14_days

knitr::kable(last_14_days, caption = paste('Safe counties across State'),
             col.names = c('County'))

```


